---
title: "Purpose2"
author: "Xinnuo Chen"
date: "2025-04-25"
output: html_document
---
Dataset with 8 normal variables, 800 observations (more cases), and fewer differential clusters.
The propensity scores used to calculate SPECKS are obtained using a logit model (previously CART). -> Better results. 

# Libraries
```{r}
library(synthpop)
library(NbClust)
library(ggplot2)
library(factoextra)
library(DescTools)
library(cluster)
library(car)
library(tidyverse)
library(clue) #label switching
library(rpart)
library(dgof)
library(dplyr)
library(tibble)
library(paletteer)
library(scales)
colour.patteer <- paletteer_c("ggthemes::Classic Orange-Blue", 80)
colours <- colour.patteer[c(23, 78, 43)]

library(tidyr)
library(mvtnorm)
```

# Functions
```{r}
# Real data generation
centroid.generation <- function(k = k, p = p, separation = separation) {
  centroids <- matrix(NA, nrow = k, ncol = p)
  centroids[1, ] <- rep(0, p)                 # first centroid at the origin 
  for (i in 2:k) {
    tries <- 0
    repeat {
      vector <- rnorm(p)                      # a random point
      vector <- vector / sqrt(sum(vector^2))  # normalize the point vector
      vector <- vector * separation           # distance of number of separations from the origin
      
      dists <- sqrt(rowSums((centroids[1:(i - 1), , drop = FALSE] -  # distances from the new point to each of the fixed points
                             matrix(vector, nrow = i - 1, ncol = p, byrow = TRUE))^2))
      if (all(dists >= separation)) {         # accept the new point if this complete the separation
        centroids[i, ] <- vector
        break
      }
      tries <- tries + 1                      # avoid infinite loop
      if (tries > 1000) stop("It has not been possible to generate sufficiently separate centroids.")
    }
  }
  return(centroids)
}

data.generation <- function(N, p, rho, separation, k){
  # position of the centroid
  centroids <- centroid.generation(k = k, p = p, separation = separation)
  
  # correlation matrix
  R <- matrix(rho, nrow = p, ncol = p)
  diag(R) <- 1
  
  # covariance matrix
  sd <- rep(1, p)
  D <- diag(sd)
  Sigma <- D %*% R %*% D  #
  
  # data generation for each cluster
  n <- round(N/k, digits = 0)
  clusters <- lapply(1:k, function(i) {
    rmvnorm(n,
            mean  = centroids[i, ],
            sigma = Sigma)
  })
  data <- do.call(rbind, clusters)
  data <- data.frame(data, group = factor(rep(1:k, each = n)))
  
  # graphic representation
  Graphic <- ggplot(data, aes(x = X1, y = X2, color = group)) +
                geom_point(size = 2) +
                scale_color_brewer(palette = "Set1") +
                labs(
                  title = "First dimention representation",
                  color = "Cluster"
                ) +
                theme_minimal(base_size = 14)
  result <- list("Data" = data[,-ncol(data)], "Graphic"=Graphic, "Clusters" = clusters)
}


# Synthetic data generation
synt.data.generation <- function(real.data, m, methods){
    syn.obj <- syn(real.data,
                       method = rep(methods, ncol(real.data)),
                       m = m,
                       print.flag = FALSE, 
                       proper = TRUE)
    synt.data <- syn.obj$syn
    specks.val <- numeric()
    if(m==1){
      names(synt.data) <- names(real.data)
      data <- rbind(
            cbind(real.data, label = 0),
            cbind(synt.data, label = 1)
          )
      specks.val <- c(specks.val, specks.correct(data = data)$Specks)
      specks.value <- data.frame(SPECKS = specks.val)
    } else {
      for(k in 1:m){
          names(synt.data[[k]]) <- names(real.data)
          
          data <- rbind(
            cbind(real.data, label = 0),
            cbind(synt.data[[k]], label = 1)
          )
          
          specks.val <- c(specks.val, specks.correct(data = data)$Specks)
          specks.value <- data.frame(SPECKS = specks.val)
      }
    }
    results <- list("Specks" = specks.value, "Syn data" = synt.data)
    return(results)
}


# K decision & clustering 
k.decision <- function(data, k.real = 0, synthetic = TRUE){ 
  if(synthetic == FALSE){
    sil <- fviz_nbclust(data, kmeans, method = "silhouette")$data
    kopt <- as.numeric(sil$clusters[which.max(sil$y)])
    
    sdata <- scale(data)
    clust <- kmeans(sdata, kopt, nstart = 25, iter.max = 100)
    
    clustering.result <- list("K optim" = kopt, "Clustering" = clust)
    return(clustering.result)
  } else {
    sil <- fviz_nbclust(data, kmeans, method = "silhouette")$data
    kopt <- as.numeric(sil$clusters[which.max(sil$y)])
    
    if(kopt == k.real){
      sdata <- scale(data)
      clust <- kmeans(sdata, kopt, nstart = 25, iter.max = 100) #resultado de clustering
      
      clustering.result <- list("K optim" = kopt, "Clustering" = clust)
      return(clustering.result)
    }
  }
}

# SPECKS CALCULATION
specks.correct <- function(data, sample=FALSE){   #### With pvalue
  if(sample==TRUE){
    data[,3] <- sample(data[,3])
  }
  
  N <- nrow(data)
  idx <- sample(1:N, round(N*0.3, digits=0))
  
  data.train <- data[-idx,]
  data.test <- data[idx,]
  
  logit_model <- glm(label ~ ., data = data.train, family = "binomial")
  
  pred.prob <- predict(logit_model, newdata = data.test, type = "response")
  
  label.vector <- data.test$label
  
  prop.score.real  <- pred.prob[label.vector == 0]
  prop.score.synt  <- pred.prob[label.vector == 1]
  
  ecdf.real  <- ecdf(prop.score.real)
  ecdf.synth <- ecdf(prop.score.synt)
  
  all.scores <- sort(unique(c(prop.score.real, prop.score.synt)))
  
  specks.value <- max(abs(ecdf.real(all.scores) - ecdf.synth(all.scores)))
  p.value <- suppressWarnings(ks.test(prop.score.real, prop.score.synt)$p)
  
  return(data.frame(
    "Specks" = specks.value, 
    "P-value" = p.value))
}

# Label switching function
label.switching <- function(km.real, km.synt){ # Label switching function
  if(!is.null(km.synt)){
      centroids <- rbind(km.real$centers, km.synt$centers)
      dist.global <- as.matrix(dist(centroids, method = "euclidean"))
      n.real <- nrow(km.real$centers) #numero de cluster en real
      n.synt <- nrow(km.synt$centers) #numero de cluster en synt
      dist <- dist.global[1:n.real, (n.real + 1) : (n.real + n.synt)]
      label.ass <- solve_LSAP(dist, maximum = FALSE) 
      
      # Reetiquetació els clústers sintètics
      km.synt$recluster <- km.real$cluster # nova variable
      
      for (i in 1:length(label.ass)) {
        km.synt$recluster[km.synt$cluster == label.ass[i]] <- i
      }
  }
  return(km.synt)
}

# Metrics function
cluster.metrics <- function(km.real, km.synt, k.real, k.synt, real.data, synt.data){
    # Nombre de cluster
    clust.num <- k.synt/k.real
    
    # Nombre d'observacions dins de clúster -> accuracy
    conf.mat <- confusionMatrix(data = as.factor(km.synt$recluster), reference = as.factor(km.real$cluster))
    accuracy <- conf.mat$overall["Accuracy"]
    
    # Distribució d'observacions en clúster -> Índex de Gini
    gini.real <- Gini(km.real$size)
    gini.synt <- Gini(km.synt$size)
    #gini <- (gini.synt)/(gini.real + 1e-6)
    gini <- abs(gini.synt - gini.real)
    
    # Distancia d'observacions a cluster -> Coeficient de silhouette (entre -1 i 1)
    sil.real <- mean(silhouette(km.real$cluster, dist(real.data))[,3])
    sil.synt <- mean(silhouette(km.synt$recluster, dist(synt.data))[,3])
    sil <- sil.synt/sil.real
    
    # Distancia mitjana entre centroides
    vars <- names(real.data)
    lreal.data <- real.data
    lreal.data$label <- as.factor(km.real$cluster)
    lsynt.data <- synt.data
    lsynt.data$label <- as.factor(km.synt$recluster)

    real.centroid <- aggregate(. ~ label,
                                data = lreal.data[, c(vars, "label")],
                                FUN = base::mean)
    synt.centroid <- aggregate(. ~ label,
                                data = lsynt.data[, c(vars, "label")],
                                FUN = base::mean)
    real.centroid <- real.centroid[order(real.centroid$label), ]
    synt.centroid <- synt.centroid[order(synt.centroid$label), ]
    
    mat.real <- as.matrix(real.centroid[, vars])
    mat.synt <- as.matrix(synt.centroid[, vars])
    distances <- sqrt(rowSums((mat.real - mat.synt)^2))
    mean.distance <- mean(distances)

    # Variancia mitjana entre centroides
    real.var <- aggregate(. ~ label,
                                data = lreal.data[, c(vars, "label")],
                                FUN = var)
    synt.var <- aggregate(. ~ label,
                                     data = lsynt.data[, c(vars, "label")],
                                     FUN = var)
    real.var <- real.var[order(real.var$label), ]
    synt.var <- synt.var[order(synt.var$label), ]
    
    mat.real.var <- as.matrix(real.var[, vars])
    mat.synt.var <- as.matrix(synt.var[, vars])
    
    dif.var <- abs(mat.real.var - mat.synt.var)
    mean.var <- mean(dif.var)
    
  result <- list("Clus.num" = clust.num, "Accuracy" = accuracy, "Gini" = gini, "Sil" = sil, "Mean distance" = mean.distance, "Mean variance" = mean.var)
  return(result)
}

plot.metrics <- function(df, specks.col = "SPECKS") {
  num_cols <- sapply(df, is.numeric)
  num_cols[specks.col] <- FALSE

  long.df <- pivot_longer(df,
                          cols      = names(num_cols)[num_cols],
                          names_to  = "Metric",
                          values_to = "Value")

  ggplot(long.df,
         aes(x = .data[[specks.col]], y = Value)) +   # tidy eval
    geom_point(alpha = 0.6) +
    geom_smooth(method = "lm", se = FALSE, colour = "red", formula = y ~ x) +
    facet_wrap(~ Metric, scales = "free_y") +
    labs(x = specks.col, y = NULL,
         title = "Relació de cada mètrica amb SPECKS") +
    theme_minimal()
}
```

# CASE 1: IDEAL CASE
-   Separation = 2sigma
-   Clust.num = 2
-   Variables = 8

## Real data generation
```{r}
my.seed <- 12345
set.seed(my.seed)

real.data <- data.generation(1000, p = 8, rho = 0, separation = 2, k = 2)$Data
head(real.data)

k.real <- k.decision(real.data, synthetic = FALSE)$`K optim`
km.real <- k.decision(real.data, synthetic = FALSE)$Clustering

```

## Synthetic data generation
```{r}
total.m <- 1000
count <- 0
m <- 0
synt.data <- list()
km.synt <- list()
k.synt <- numeric()
speck <- numeric()
while(m < total.m){
  set.seed(m + count)
  count <- count + 1
  syn   <- synt.data.generation(real.data, m = 1, methods = "cart")
  clust <- k.decision(syn$`Syn data`, k.real = k.real, synthetic = TRUE)
  if(!is.null(clust)){
    m <- m + 1
    synt.data[[m]] <- syn$`Syn data`
    km.synt[[m]] <- clust$Clustering
    k.synt[m] <- clust$`K optim`
    speck[m] <- syn$Specks$SPECKS
  }
}
```


## Metrics
```{r}
vars <- names(real.data)
gini.real <- Gini(km.real$size)
sil.real <- mean(silhouette(km.real$cluster, dist(real.data))[,3])
lreal.data <- real.data
lreal.data$label <- as.factor(km.real$cluster)

ls.km.synt <- lapply(km.synt, label.switching, km.real) #label switching

  ## Gini
  gini.syn <- sapply(km.synt, function(km) Gini(km$size))
  
  sil.synt <- numeric(m)      
  mean.distance <- numeric(m)
  mean.var <- numeric(m)
  for (i in seq_along(km.synt)){
    lsynt.data <- synt.data[[i]]
    lsynt.data$label <- factor(ls.km.synt[[i]]$recluster)
    
    ## Silhouette
    sil.synt[i] <- mean(silhouette(km.synt[[i]]$cluster, dist(synt.data[[i]]))[,3])
    
    ## Average distance between centroids
    real.centroid <- aggregate(. ~ label,
                                data = lreal.data[, c(vars, "label")],
                                FUN = base::mean)
    synt.centroid <- aggregate(. ~ label,
                                data = lsynt.data[, c(vars, "label")],
                                FUN = base::mean)
    real.centroid <- real.centroid[order(real.centroid$label), ]
    synt.centroid <- synt.centroid[order(synt.centroid$label), ]
        
    mat.real <- as.matrix(real.centroid[, vars])
    mat.synt <- as.matrix(synt.centroid[, vars])
    distances <- sqrt(rowSums((mat.real - mat.synt)^2))
    mean.distance[i] <- mean(distances)
    
    ## Mean variance between centroids
    real.var <- aggregate(. ~ label,
                          data = lreal.data[, c(vars, "label")],
                          FUN = var)
    synt.var <- aggregate(. ~ label,
                          data = lsynt.data[, c(vars, "label")],
                          FUN = var)
    real.var <- real.var[order(real.var$label), ]
    synt.var <- synt.var[order(synt.var$label), ]
    
    mat.real.var <- as.matrix(real.var[, vars])
    mat.synt.var <- as.matrix(synt.var[, vars])
    
    dif.var <- abs(mat.real.var - mat.synt.var)
    mean.var[i] <- mean(dif.var)
  }

result_case1 <- data.frame(Specks = speck,
                           Gini.real = rep(gini.real, m),
                           Gini.synt = gini.syn,
                           Sil.real = rep(sil.real, m),
                           Sil.synt = sil.synt,
                           Mean.distance = mean.distance,
                           Mean.var = mean.var)
result_case1 <- result_case1 %>%
  mutate(Diff.gini = Gini.real - Gini.synt) %>%
  mutate(Diff.sil = Sil.real - Sil.synt)
```

## Graphics
```{r}
metrics <- c("Diff.sil", "Diff.gini", "Mean.distance", "Mean.var")

graphics1 <- list(length(metrics))

for (i in seq_along(metrics)) {
  graphics1[[i]] <- ggplot(result_case1, aes_string(x = "Specks", y = metrics[i])) +
    geom_point(color = "#3182bd", size = 1) +
    geom_smooth(method = "lm", se = TRUE, color = "#e6550d") +
    theme_minimal(base_size = 14)
}
print(graphics1)

# library(patchwork)
layout1 <- (graphics1[[1]] | graphics1[[2]]) /
          (graphics1[[3]] | graphics1[[4]])

print(layout1)

library(ggplot2)
ggsave(
  filename = "specks_metrics_case1.png",
  plot     = layout1,
  width    = 12,
  height   = 6,
  dpi      = 300
)
```


# CASE 2: PROBLEMATIC CAS
-   Separation = 10sigma
-   Clust.num = 4
-   Variables = 10

## Real data generation
```{r}
my.seed <- 12345
set.seed(my.seed)

real.data <- data.generation(1000, p = 10, rho = 0, separation = 10, k = 4)$Data
head(real.data)

k.real <- k.decision(real.data, synthetic = FALSE)$`K optim`
km.real <- k.decision(real.data, synthetic = FALSE)$Clustering

```

## Synthetic data generation
```{r}
total.m <- 1000
count <- 0
m <- 0
synt.data <- list()
km.synt <- list()
k.synt <- numeric()
speck <- numeric()
while(m < total.m){
  set.seed(m + count)
  count <- count + 1
  syn   <- synt.data.generation(real.data, m = 1, methods = "cart")
  clust <- k.decision(syn$`Syn data`, k.real = k.real, synthetic = TRUE)
  if(!is.null(clust)){
    m <- m + 1
    synt.data[[m]] <- syn$`Syn data`
    km.synt[[m]] <- clust$Clustering
    k.synt[m] <- clust$`K optim`
    speck[m] <- syn$Specks$SPECKS
  }
}
```

## Metrics
```{r}
vars <- names(real.data)
gini.real <- Gini(km.real$size)
sil.real <- mean(silhouette(km.real$cluster, dist(real.data))[,3])
lreal.data <- real.data
lreal.data$label <- as.factor(km.real$cluster)

ls.km.synt <- lapply(km.synt, label.switching, km.real) #label switching

  ## Gini
  gini.syn <- sapply(km.synt, function(km) Gini(km$size))
  
  sil.synt <- numeric(m)      
  mean.distance <- numeric(m)
  mean.var <- numeric(m)
  for (i in seq_along(km.synt)){
    lsynt.data <- synt.data[[i]]
    lsynt.data$label <- factor(ls.km.synt[[i]]$recluster)
    
    ## Silhouette
    sil.synt[i] <- mean(silhouette(km.synt[[i]]$cluster, dist(synt.data[[i]]))[,3])
    
    ## Average distance between centroids
    real.centroid <- aggregate(. ~ label,
                                data = lreal.data[, c(vars, "label")],
                                FUN = base::mean)
    synt.centroid <- aggregate(. ~ label,
                                data = lsynt.data[, c(vars, "label")],
                                FUN = base::mean)
    real.centroid <- real.centroid[order(real.centroid$label), ]
    synt.centroid <- synt.centroid[order(synt.centroid$label), ]
        
    mat.real <- as.matrix(real.centroid[, vars])
    mat.synt <- as.matrix(synt.centroid[, vars])
    distances <- sqrt(rowSums((mat.real - mat.synt)^2))
    mean.distance[i] <- mean(distances)
    
    ## Average variance between centroids
    real.var <- aggregate(. ~ label,
                          data = lreal.data[, c(vars, "label")],
                          FUN = var)
    synt.var <- aggregate(. ~ label,
                          data = lsynt.data[, c(vars, "label")],
                          FUN = var)
    real.var <- real.var[order(real.var$label), ]
    synt.var <- synt.var[order(synt.var$label), ]
    
    mat.real.var <- as.matrix(real.var[, vars])
    mat.synt.var <- as.matrix(synt.var[, vars])
    
    dif.var <- abs(mat.real.var - mat.synt.var)
    mean.var[i] <- mean(dif.var)
  }

result_case2 <- data.frame(Specks = speck,
                           Gini.real = rep(gini.real, m),
                           Gini.synt = gini.syn,
                           Sil.real = rep(sil.real, m),
                           Sil.synt = sil.synt,
                           Mean.distance = mean.distance,
                           Mean.var = mean.var)
result_case2 <- result_case2 %>%
  mutate(Diff.gini = Gini.real - Gini.synt) %>%
  mutate(Diff.sil = Sil.real - Sil.synt)
```

## Graphics
```{r}
metrics <- c("Diff.sil", "Diff.gini", "Mean.distance", "Mean.var")

graphics2 <- list(length(metrics))

for (i in seq_along(metrics)) {
  graphics2[[i]] <- ggplot(result_case2, aes_string(x = "Specks", y = metrics[i])) +
    geom_point(color = "#3182bd", size = 1) +
    geom_smooth(method = "lm", se = TRUE, color = "#e6550d") +
    theme_minimal(base_size = 14)
}
print(graphics2)

library(patchwork)
layout2 <- (graphics2[[1]] | graphics2[[2]]) /
          (graphics2[[3]] | graphics2[[4]])

print(layout2)

library(ggplot2)
ggsave(
  filename = "specks_metrics_case2.png",
  plot     = layout2,
  width    = 12,
  height   = 6,
  dpi      = 300
)
```

# CASE 3: Others
-   Separation = 2sigma
-   Clust.num = 4
-   Variables = 2

## Real data generation
```{r}
my.seed <- 12345
set.seed(my.seed)

real.data <- data.generation(1000, p = 2, rho = 0, separation = 2, k = 4)$Data
head(real.data)

k.real <- k.decision(real.data, synthetic = FALSE)$`K optim`
km.real <- k.decision(real.data, synthetic = FALSE)$Clustering

```

## Synthetic data generation
```{r}
total.m <- 1000
count <- 0
m <- 0
synt.data <- list()
km.synt <- list()
k.synt <- numeric()
speck <- numeric()
while(m < total.m){
  set.seed(m + count)
  count <- count + 1
  syn   <- synt.data.generation(real.data, m = 1, methods = "cart")
  clust <- k.decision(syn$`Syn data`, k.real = k.real, synthetic = TRUE)
  if(!is.null(clust)){
    m <- m + 1
    synt.data[[m]] <- syn$`Syn data`
    km.synt[[m]] <- clust$Clustering
    k.synt[m] <- clust$`K optim`
    speck[m] <- syn$Specks$SPECKS
  }
}
```

## Metrics
```{r}
vars <- names(real.data)
gini.real <- Gini(km.real$size)
sil.real <- mean(silhouette(km.real$cluster, dist(real.data))[,3])
lreal.data <- real.data
lreal.data$label <- as.factor(km.real$cluster)

ls.km.synt <- lapply(km.synt, label.switching, km.real) #label switching

  ## Gini
  gini.syn <- sapply(km.synt, function(km) Gini(km$size))
  
  sil.synt <- numeric(m)      
  mean.distance <- numeric(m)
  mean.var <- numeric(m)
  for (i in seq_along(km.synt)){
    lsynt.data <- synt.data[[i]]
    lsynt.data$label <- factor(ls.km.synt[[i]]$recluster)
    
    ## Silhouette
    sil.synt[i] <- mean(silhouette(km.synt[[i]]$cluster, dist(synt.data[[i]]))[,3])
    
    ## Average distance between centroids
    real.centroid <- aggregate(. ~ label,
                                data = lreal.data[, c(vars, "label")],
                                FUN = base::mean)
    synt.centroid <- aggregate(. ~ label,
                                data = lsynt.data[, c(vars, "label")],
                                FUN = base::mean)
    real.centroid <- real.centroid[order(real.centroid$label), ]
    synt.centroid <- synt.centroid[order(synt.centroid$label), ]
        
    mat.real <- as.matrix(real.centroid[, vars])
    mat.synt <- as.matrix(synt.centroid[, vars])
    distances <- sqrt(rowSums((mat.real - mat.synt)^2))
    mean.distance[i] <- mean(distances)
    
    ## Average variance between centroids
    real.var <- aggregate(. ~ label,
                          data = lreal.data[, c(vars, "label")],
                          FUN = var)
    synt.var <- aggregate(. ~ label,
                          data = lsynt.data[, c(vars, "label")],
                          FUN = var)
    real.var <- real.var[order(real.var$label), ]
    synt.var <- synt.var[order(synt.var$label), ]
    
    mat.real.var <- as.matrix(real.var[, vars])
    mat.synt.var <- as.matrix(synt.var[, vars])
    
    dif.var <- abs(mat.real.var - mat.synt.var)
    mean.var[i] <- mean(dif.var)
  }

result_case3 <- data.frame(Specks = speck,
                           Gini.real = rep(gini.real, m),
                           Gini.synt = gini.syn,
                           Sil.real = rep(sil.real, m),
                           Sil.synt = sil.synt,
                           Mean.distance = mean.distance,
                           Mean.var = mean.var)
result_case3 <- result_case3 %>%
  mutate(Diff.gini = Gini.real - Gini.synt) %>%
  mutate(Diff.sil = Sil.real - Sil.synt)
```

## Graphics
```{r}
metrics <- c("Diff.sil", "Diff.gini", "Mean.distance", "Mean.var")

graphics3 <- list(length(metrics))

for (i in seq_along(metrics)) {
  graphics3[[i]] <- ggplot(result_case3, aes_string(x = "Specks", y = metrics[i])) +
    geom_point(color = "#3182bd", size = 1) +
    geom_smooth(method = "lm", se = TRUE, color = "#e6550d") +
    theme_minimal(base_size = 14)
}
print(graphics3)

library(patchwork)
layout3 <- (graphics3[[1]] | graphics3[[2]]) /
          (graphics3[[3]] | graphics3[[4]])

print(layout3)

library(ggplot2)
ggsave(
  filename = "specks_metrics_case2.png",
  plot     = layout,
  width    = 12,
  height   = 6,
  dpi      = 300
)
```
